---
title: "Cor_by_GeneSetSize"
author: "Avery Williams"
date: '2022-08-13'
output: html_document
---

> Evaluating differences in cell line/pdx performance in correlating to origin tissue depending on size of gene set, whether highly tumor purity-correlated genes are included/excluded, and whether gene rankings by most variable are based on TCGA or GTEx (i.e., whether they are ranked by variability across tumor types or across control tissues)

```{r source load}
source("/Users/avery/Desktop/work/R Projects/modelselection/scripts/functions/tissuecormatch.R") # change to where you have this function script saved
source("/Users/avery/Desktop/work/R Projects/modelselection/scripts/functions/tissuecormatch_median.R")
```

# I copied full (63856 genes) rses and tpm from my recount3 download folder into a /Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/63856/ folder to be in the same root location as the subsets, in ../63856/tpm and ../63856/rse folders

# run correlation fun
```{r run fun}
for(i in c(100, 1000, 5000, 10000)){
  sub <- i
  for(i in c("purinc", "purexc")){
    pur <- i
    for(i in c("gtex", "tcga")){
      var <- i
      
      tissuecormatch(rootfilepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", row_sources = c("bone_c", "ccle", "hpa_c"), col_sources = c("tcga", "gtex", "hpa_t", "bone_n", "bone_t"), sub = sub, pur = pur, var = var, app_by = "disease", save = TRUE, savefilepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/")
    }
  }
}
# making object for ALL model sources
tissuecormatch(rootfilepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", row_sources = c("bone_c", "ccle", "hpa_c", "pdx"), col_sources = c("tcga", "gtex", "hpa_t", "bone_n", "bone_t"), sub = 63856, app_by = "disease", save = TRUE, savefilepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/")

# now just cell lines
tissuecormatch(rootfilepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", row_sources = c("bone_c", "ccle", "hpa_c", "pdx"), col_sources = c("tcga", "gtex", "hpa_t", "bone_n", "bone_t"), sub = 63856, app_by = "disease", save = TRUE, savefilepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/")

rm(sub, pur, var)
```

# visualize correlation in a table
```{r correlation table}
percent_map_to_origin <- c()
id <- c()
for(i in c(100, 1000, 5000, 10000)){
  sub <- i
  for(i in c("tcga", "gtex")){
    var <- i
    for(i in c("purinc", "purexc")){
       cor_temp <- readRDS(paste("/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", sub, "/", i, "/", var, "var/corres/", "cor_dismatch_", i, "_", var, "var_res_", sub, ".rds", sep = ""))[["dfo_summary"]]
      percent_map_to_origin <- c(percent_map_to_origin, cor_temp$same_as_origin/(cor_temp$different + cor_temp$same_as_origin))
      id <- c(id, paste(sub,var,i, sep = "_"))
    }
  }
}
View(cbind(id, percent_map_to_origin))
rm(i, id, sub, var, cor_temp)
```

# Plot how % mapping to origin changes between subsets
```{r}
corr_table <- data.frame(genes_included = c(rep(100,4),rep(1000,4),rep(5000,4),rep(10000,4)), percent_map_to_origin = percent_map_to_origin, Variation = rep(c("TCGA", "TCGA", "GTEx", "GTEx"),4), Purity = rep(c("Included", "Excluded"),8))
library(ggplot2)
ggplot(corr_table, aes(x = genes_included, y = percent_map_to_origin, color = Variation, linetype = Purity)) + geom_line() + geom_point() + geom_hline(aes(yintercept=0.237579), color="darkslategray", linetype="dashed") + labs(x = "Number of Genes Included", y = "Percent of Cell Lines That Correlate Highest to Origin Tissue")
ggsave("/Users/avery/Desktop/work/R Projects/modelselection/plots/percentmaptoorigin.png")
rm(percent_map_to_origin, corr_table)
```

# run function to make a bunch of objects including/excluding purity, of various gene set sizes (5K, 10K, full), and variation source as TCGA or GTEx
```{r run function}
# TO VIEW AS A DF
tissuecormatch_med_res <- data.frame(
  oneH_I_T = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 100, "purinc", "tcga"),
  oneH_E_T = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 100, "purexc", "tcga"),
  oneH_I_G = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 100, "purinc", "gtex"),
  oneH_E_G = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 100, "purexc", "gtex"),
  oneK_I_T = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 1000, "purinc", "tcga"),
  oneK_E_T = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 1000, "purexc", "tcga"),
  oneK_I_G = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 1000, "purinc", "gtex"),
  oneK_E_G = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 1000, "purexc", "gtex"),
  fiveK_I_T = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 5000, "purinc", "tcga"),
  fiveK_E_T = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 5000, "purexc", "tcga"),
  fiveK_I_G = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 5000, "purinc", "gtex"),
  fiveK_E_G = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 5000, "purexc", "gtex"),
  tenK_I_T = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 10000, "purinc", "tcga"),
  tenK_E_T = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 10000, "purexc", "tcga"),
  tenK_I_G = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 10000, "purinc", "gtex"),
  tenK_E_G = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 10000, "purexc", "gtex"),
  ALL = tissuecormatch_median(filepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", 63856)
)

colnames(tissuecormatch_med_res) <- c("oneH_I_T", "oneH_E_T", "oneH_I_G", "oneH_E_G", "oneK_I_T", "oneK_E_T", "oneK_I_G", "oneK_E_G", "fiveK_I_T", "fiveK_E_T", "fiveK_I_G", "fiveK_E_G", "tenK_I_T", "tenK_E_T", "tenK_I_G", "tenK_E_G", "ALL")
```

# summary line plot of median of each group size across all tissues
```{r}
corr_median_lp <- data.frame(gene_set_size = rep(c(100,1000,5000,10000), 4), Purity = c(rep("Included", 4), rep("Excluded", 4), rep("Included", 4), rep("Excluded", 4)), Variation = c(rep("TCGA", 8), rep("GTEx", 8)), median_cor_across_tissues = c(median(tissuecormatch_med_res$oneH_I_T), median(tissuecormatch_med_res$oneK_I_T), median(tissuecormatch_med_res$fiveK_I_T), median(tissuecormatch_med_res$tenK_I_T), median(tissuecormatch_med_res$oneH_E_T), median(tissuecormatch_med_res$oneK_E_T), median(tissuecormatch_med_res$fiveK_E_T), median(tissuecormatch_med_res$tenK_E_T), median(tissuecormatch_med_res$oneH_I_G), median(tissuecormatch_med_res$oneK_I_G), median(tissuecormatch_med_res$fiveK_I_G), median(tissuecormatch_med_res$tenK_I_G), median(tissuecormatch_med_res$oneH_E_G), median(tissuecormatch_med_res$oneK_E_G), median(tissuecormatch_med_res$fiveK_E_G), median(tissuecormatch_med_res$tenK_E_G)))

library(ggplot2)
ggplot(corr_median_lp, aes(x = gene_set_size, y = median_cor_across_tissues, color = Variation, linetype = Purity)) + geom_line() + geom_point() + labs(x = "Number of Genes Included", y = "Median Correlation of Cell Lines to Their Origin Tissue") + geom_hline(aes(yintercept=median(tissuecormatch_med_res$ALL)), color="darkslategray", linetype="dashed")

ggsave("/Users/avery/Desktop/work/R Projects/modelselection/plots/medcortoorigin.png")
rm(corr_median_lp)
```

## I wonder how this plot is different if I take the raw median and not a median of medians
```{r}
# what genes to subset to
subset_gene_lists <- list()
for(i in c(100, 1000, 5000, 10000)){
  num <- i
  for(i in c("purinc", "purexc")){
    pur <- i
    for(i in c("tcgavar", "gtexvar")){
      var <- i
      subset_gene_lists[[paste(num)]][[pur]][[var]] <- rownames(readRDS(paste("/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", num, "/", pur, "/", var, "/tpm/bone_c_tpm_sub", num, "_", pur, "_", var, ".rds", sep = "")))
    }
  }
}

# for each set, take the median
library(matrixStats)

subset_medians <- c()
subsets <- c()
for(i in c(100, 1000, 5000, 10000)){
  num <- i
  for(i in c("purinc", "purexc")){
    pur <- i
    for(i in c("tcga", "gtex")){
      var <- i
      
      subset_medians <- c(subset_medians, median(rowMedians(tissuecormatch(rootfilepath = "/Users/avery/Desktop/work/R Projects/modelselection/data/rse_subsets/", row_sources = c("bone_c", "ccle", "hpa_c"), col_sources = c("tcga", "gtex", "hpa_t", "bone_n", "bone_t"), sub = num, pur = pur, var = var, app = FALSE, save = FALSE)[["fullcor"]])))
      
      subsets <- c(subsets, paste(num, pur, var, sep = "_"))
    }
  }
}
```

```{r}
fullcor <- cor(cbind(gtex_tpm, tcga_tpm, bone_n_tpm, bone_t_tpm, hpa_t_tpm), cbind(ccle_tpm, pdx_tpm, hpa_c_tpm, bone_c_tpm), method = "spearman") # this object takes a while, only build if not already in env/exported
saveRDS(fullcor, "~/Desktop/work/R Projects/modelselection/data/fullcor.rds") # so it doesn't need to be built again

subset_medians <- c(subset_medians, median(rowMedians(fullcor)))
subsets <- c(subsets, "full")

subset_table <- data.frame(category = subsets, median = subset_medians)
```

## make a line plot with it
```{r}
subset_table_lp <- data.frame(name = subsets[-17], median_cor_across_tissues = subset_medians[-17], Variation = NA, Purity = NA, gene_set_size = c(rep(100,4), rep(1000,4), rep(5000,4), rep(10000,4)))

subset_table_lp[grep("tcga", subset_table_lp$name),]$Variation <- "tcga"
subset_table_lp[grep("gtex", subset_table_lp$name),]$Variation <- "gtex"

subset_table_lp[grep("inc", subset_table_lp$name),]$Purity <- "inc"
subset_table_lp[grep("exc", subset_table_lp$name),]$Purity <- "exc"

# plot
library(ggplot2)
ggplot(subset_table_lp, aes(x = gene_set_size, y = median_cor_across_tissues, color = Variation, linetype = Purity)) + geom_line() + geom_point() + labs(x = "Number of Genes Included", y = "Median Correlation of Cell Lines to Their Origin Tissue") + geom_hline(aes(yintercept=subset_medians[17]), color="darkslategray", linetype="dashed")

ggsave("/Users/avery/Desktop/work/R Projects/modelselection/plots/medcortoorigin_fromfullcor.png") # this is used in final over other one
```

## make bubble plot for individual tissues
### these are not used in final version
```{r}
# lengthwise dataframe
tissuecormatch_med_cor_long <- c()
for(i in 1:ncol(tissuecormatch_med_res)){
  tissuecormatch_med_cor_long <- c(tissuecormatch_med_cor_long, tissuecormatch_med_res[,i])
}

tissuecormatch_med_bp <- data.frame(cor = tissuecormatch_med_cor_long, Cancer = rep(substr(rownames(tissuecormatch_med_res),1,nchar(rownames(tissuecormatch_med_res))-7), ncol(tissuecormatch_med_res)), Subset = c(rep(100,4 * nrow(tissuecormatch_med_res)), rep(1000,4 * nrow(tissuecormatch_med_res)), rep(5000,4 * nrow(tissuecormatch_med_res)), rep(10000,4 * nrow(tissuecormatch_med_res)), rep(63856, nrow(tissuecormatch_med_res))), Purity = c(rep(c(rep("inc",nrow(tissuecormatch_med_res)),rep("exc",nrow(tissuecormatch_med_res))),8), rep("All", nrow(tissuecormatch_med_res))), Variance = c(rep(c(rep("TCGA",nrow(tissuecormatch_med_res)*2),rep("GTEx",nrow(tissuecormatch_med_res)*2)),4), rep("All", nrow(tissuecormatch_med_res))))

ggplot(tissuecormatch_med_bp, aes(x = Cancer, y = cor, size = Subset, color = Variance, shape = Purity)) + geom_count() + theme(axis.text.x = element_text(angle = 45, hjust=1)) + labs(y = "Median Correlation of Cell Lines to Matched Tissue")

# split into 2 to lessen complexity
ggplot(tissuecormatch_med_bp[tissuecormatch_med_bp$Purity != "exc",], aes(x = Cancer, y = cor, size = Subset, color = Variance)) + geom_count() + theme(axis.text.x = element_text(angle = 45, hjust=1)) + labs(y = "Median Correlation of Cell Lines to Matched Tissue") + ggtitle("Including Tumor Purity-Correlated Genes") + ylim(0,1)
ggsave("/Users/avery/Desktop/work/R Projects/modelselection/plots/medcortoorigin_tissplit_incpur.png")

ggplot(tissuecormatch_med_bp[tissuecormatch_med_bp$Purity != "inc",], aes(x = Cancer, y = cor, size = Subset, color = Variance)) + geom_count() + theme(axis.text.x = element_text(angle = 45, hjust=1)) + labs(y = "Median Correlation of Cell Lines to Matched Tissue") + ggtitle("Excluding Tumor Purity-Correlated Genes") + ylim(0,1)
ggsave("/Users/avery/Desktop/work/R Projects/modelselection/plots/medcortoorigin_tissplit_excpur.png")
```

# redo bubble plot as a heatmap
## trying including subset size as a row variable with cancer type, but becomes too long, so sticking with tissuecormatch_med_res object still
```{r}
for(i in c("I_T", "E_T", "I_G", "E_G")){
  colsub <- tissuecormatch_med_res[,grep(i, colnames(tissuecormatch_med_res))]
  temp <- c(colsub[,1], colsub[,2], colsub[,3], colsub[,4])
  if(which(c("I_T", "E_T", "I_G", "E_G") == i) == 1){
    tissuecormatch_med_resedit <- (data.frame(col1 = temp))
  } else if(which(c("I_T", "E_T", "I_G", "E_G") == i) != 1){
    tissuecormatch_med_resedit <- cbind(tissuecormatch_med_resedit, temp)
  }
}
colnames(tissuecormatch_med_resedit) <- c("I_T", "E_T", "I_G", "E_G")
rownames(tissuecormatch_med_resedit) <- c(paste(substr(rownames(tissuecormatch_med_res),1, nchar(rownames(tissuecormatch_med_res))-7), "_100", sep = ""), paste(substr(rownames(tissuecormatch_med_res),1, nchar(rownames(tissuecormatch_med_res))-7), "_1000", sep = ""), paste(substr(rownames(tissuecormatch_med_res),1, nchar(rownames(tissuecormatch_med_res))-7), "_5000", sep = ""), paste(substr(rownames(tissuecormatch_med_res),1, nchar(rownames(tissuecormatch_med_res))-7), "_10000", sep = ""))

# add identifier columns/rows to make annotations easier
tissuecormatch_med_resedit$setsize <- c(rep(100, length(rownames(tissuecormatch_med_res))), rep(1000, length(rownames(tissuecormatch_med_res))), rep(5000, length(rownames(tissuecormatch_med_res))), rep(10000, length(rownames(tissuecormatch_med_res))))

tissuecormatch_med_resedit$cancer <- rep(substr(rownames(tissuecormatch_med_res),1, nchar(rownames(tissuecormatch_med_res))-7),4)

# clean up
rm(colsub, temp)
```

## adding annotation rows
```{r}
tissuecormatch_med_res_annorow <- tissuecormatch_med_res
temp <- data.frame(subsetsize=c(rep(100,4), rep(1000,4), rep(5000,4), rep(10000,4), 63856), tumorpurity= c(rep(c("Included", "Excluded"),8), "Included"), variance = c(rep(c("TCGA", "TCGA", "GTEx", "GTEx"),4), "All"))
rownames(temp) <- colnames(tissuecormatch_med_res_annorow)
tissuecormatch_med_res_annorow <- rbind(tissuecormatch_med_res_annorow, t(temp))
rownames(tissuecormatch_med_res_annorow) <- c(substr(rownames(tissuecormatch_med_res),1, nchar(rownames(tissuecormatch_med_res))-7), "Subset Size", "Tumor Purity", "Variance") # adjust to remove "Cancer"
```

```{r}
library(ComplexHeatmap)

# column annotations
column_ha <- HeatmapAnnotation(df = as.data.frame(t(tissuecormatch_med_res_annorow[23:25,])),
                               col = list("Subset Size" = c("  100" = "#C86991", " 1000" = "#FEE440", " 5000" = "#328651", "10000" = "#00BBF9", "63856" = "#9B5DE5"),
                                          "Tumor Purity" = c("Included" = "#3C8EA9", "Excluded" = "#F15BB5"),
                                          "Variance" = c("TCGA" = "#00F5D4", "GTEx" = "#77267D", "All" = "#F5743D"))) # idk why the spaces are/need to be in sub size but they do

png(file="/Users/avery/Desktop/work/R Projects/modelselection/plots/genesetsize_hm_indcancers.png", width = 700, height = 500)
hm_obj <- Heatmap(as.matrix(tissuecormatch_med_res), name = "Median Correlation", row_title = "Cancer Type", row_labels = substr(rownames(tissuecormatch_med_res),1, nchar(rownames(tissuecormatch_med_res))-7), row_title_side = "right", column_labels = tissuecormatch_med_res_annorow[23,], top_annotation = column_ha, column_title = "Gene Subset Size", column_title_side = "top", column_names_side = "top", column_dend_side = "bottom", column_split = c(rep("", length(colnames(tissuecormatch_med_res))-1), 63856)) 
draw(hm_obj)
dev.off()

 ht <- Heatmap(...)
 draw(ht)
 dev.off()

```