---
title: "DataDownload"
author: "Avery Williams"
date: '2022-08-10'
output: html_document
---


> utilizing a function to pull RSEs from Recount3, make adjustments/filters to remove bad data, and save the RSEs as well as raw counts and TPM objects.

# required packages
```{r load packages}
suppressPackageStartupMessages({
  library(recount)
  library(recount3)
  library(readr)
  source("../scripts/functions/rec3_rse_download.R") # change to where you have this function script saved
  library(foreach)
  library(parallel)
  library(doParallel)
  library(tidyverse)
  library(vroom)
})
```

# utilize functions in rec3_rse_download.R

## All non-TCGA/GTEx Studies
getSRAproj function accepts available projects ("ccle", "hpa_c", "hpa_t", "pdx", "bone_t", "bone_n", "bone_c") and filepath to cellosaurus file (if repo was cloned, should be at ../data/problematic_celllines.csv) and pulls RSE's, fixes mislabels, calculates TPM, and saves whole RSE's as .rds files in 'rse' folder and RDS files for TPM and raw counts matrices in 'tpm' and 'raw_counts' folders, respectively
```{r}
startTime <- Sys.time()
for(i in c("ccle", "hpa_c", "hpa_t", "pdx", "bone_t", "bone_n", "bone_c")){
 getSRAproj(project = i, cellosaurusfilepath = "../data/problematic_celllines.csv", savefilepath = "../data/recount3/") # will place full rse (with calc TPM), rawcounts, and tpm objects in destination folder
}
endTime <- Sys.time()
print(endTime - startTime) # Time difference of 5.324986 mins

```

## TCGA/GTEx Studies
Due to large project sizes and sub-projects by tissue for TCGA and GTEx: getTCGArse() and getGTEXrse() are used to pull RSE's, run TPM calculation, and save whole RSE's as .rds files in 'rse' folder while running in parallel. Then getRawCounts() and getTPM() are used to run in parallel pulling in each RSE for GTEx and TCGA and then bind TPM and raw counts matrices and save them in 'tpm' and 'raw_counts' folders, respectively

```{r}
#GTEx
projects <- c("BRAIN", "SKIN", "ESOPHAGUS", "BLOOD", "BLOOD_VESSEL", "ADIPOSE_TISSUE", "HEART", "MUSCLE", "COLON", 
            "THYROID", "NERVE", "LUNG", "BREAST", "TESTIS", "STOMACH", "PANCREAS", "PITUITARY", "ADRENAL_GLAND", 
            "PROSTATE", "SPLEEN", "LIVER", "OVARY", "SMALL_INTESTINE", "SALIVARY_GLAND", "VAGINA", "UTERUS", 
            "KIDNEY", "BLADDER", "CERVIX_UTERI", "FALLOPIAN_TUBE")

cl <- makeCluster(8)
registerDoParallel(cl, cores = 8)
startTime <- Sys.time()

for( i in 1:length(projects)){
  getGTEXrse(project = projects[i], savefilepath = "../data/recount3/")
}

getRawCounts(project = "gtex", savefilepath = "../data/recount3/rse/gtex_tissue/") %>% saveRDS(file = "../data/recount3/raw_counts/gtex_rawcounts.rds")
getTPM(project = "gtex", savefilepath = "../data/recount3/rse/gtex_tissue/") %>% saveRDS(file = "../data/recount3/tpm/gtex_tpm.rds")

endTime <- Sys.time()
print(endTime - startTime) # Time difference of 1.5
stopCluster(cl) 
gc()
```
 
 
```{r}
 #TCGA
 projects <- c("BRCA","KIRC","LUAD","UCEC","THCA","PRAD","LUSC","HNSC","COAD","LGG","SKCM","LAML", "STAD","BLCA","OV","LIHC",
            "KIRP","CESC","SARC","ESCA","PCPG","PAAD","READ","GBM","TGCT","THYM","KICH","MESO","UVM","ACC","UCS","DLBC","CHOL")
 
 cl <- makeCluster(8)
registerDoParallel(cl, cores = 8)
startTime <- Sys.time()

for( i in 1:length(projects)){
  getTCGArse(project = projects[i], savefilepath = "../data/recount3/")
}

getRawCounts(project = "tcga", savefilepath = "../data/recount3/rse/tcga_tissue/") %>% saveRDS(file = "../data/recount3/raw_counts/tcga_rawcounts.rds")
getTPM(project = "tcga", savefilepath = "../data/recount3/rse/tcga_tissue/") %>% saveRDS(file = "../data/recount3/tpm/tcga_tpm.rds")
endTime <- Sys.time()
print(endTime - startTime) # Time difference of 44 mins
stopCluster(cl) 
gc()
```

```{r}
sessionInfo()
```
