# making a master matrix with sample names (specifically their colnames for TPM), tissue/cell line type, and disease source
# PREREQUISITES: load barebones_cellline.R function to make cell lines cross-comparable despite naming differences, load all RSE objects
ORDER: TCGA, GTEx, HPA Tissues, CCLE, HPA Cell Lines, PDX, BONE Tumor, BONE "Non-Diseased", BONE Cell Lines

# load packages
```{r load packages}
library(stringr)
library(stringi)
library(tidyverse)
source("../scripts/functions/barebones_cellline.R") # change to where you have this function script saved
library(foreach)
library(parallel)
library(doParallel)
```

#load in RSE's
```{r}
#read in all RSE's that are non-GTEx/TCGA
#function adapted from "jasondubois/sportfish"
ReadRDSFiles("../data/recount3/rse", envir = .GlobalEnv)

#read in TCGA and GTEx sub-projects (tissues) RSE's, pull metadata, and rbind by larger project (TCGA and GTEx)
cl <- makeCluster(8)
registerDoParallel(cl, cores = 8)
tcga_meta <- getMetadata("../data/recount3/rse/tcga_tissue")

gtex_meta <- getMetadata("../data/recount3/rse/gtex_tissue")
stopCluster(cl) 
```
# tissue source for HPA tissues
```{r write out hpa tissues}
# hpa metadata doesn't specify general tissue source
hpa_tissues <- c("Skin", "Skin", "Brain", "Brain", "Bone", "Lung", "Lung", "Colon", "Colon", "Kidney", "Kidney", "Cervix", "Cervix", "Liver", "Liver", "Breast", "Breast", "Prostate", "Prostate", "Bladder", "Bladder", "Brain", "Brain")
```

# build metadata data frame
tcga and gtex
```{r sample master metadata}
tcga_meta_adj <- tcga_meta %>% 
  mutate(., source = "TCGA", origin = "Tumor Tissue", rse_samp_name = rownames(tcga_meta), tissue = `tcga.gdc_cases.project.primary_site`, cell_line = NA, 
         disease = `tcga.gdc_cases.project.name`, sequencer = `tcga.gdc_platform`, sex = str_to_title(`tcga.gdc_cases.demographic.gender`), 
         race = str_to_title(`tcga.gdc_cases.demographic.race`), readlength = `recount_qc.star.average_input_read_length`, .keep = "used") %>% replace_na(list(sex = "Not Reported", race = "Not reported")) %>% #replace NAs with 'not reported'
  dplyr::select(source, origin, rse_samp_name, tissue, cell_line, disease, sequencer, sex, race, readlength)

 gtex_meta_adj <- gtex_meta %>% 
   mutate(., source = "GTEx", origin = "NonDisease Tissue", rse_samp_name = rownames(gtex_meta), tissue = `gtex.smts`, cell_line = NA, disease = paste("NonDisease", `gtex.smts`), 
          sequencer = "Illumina HiSeq", sex = ifelse(gtex.sex == 1, "Male", ifelse(gtex.sex == 2, "Female", "Not Reported")), race = "Not Reported", 
          readlength = `recount_qc.star.average_input_read_length`, .keep = "used") %>% 
   dplyr::select(source, origin, rse_samp_name, tissue, cell_line, disease, sequencer, sex, race, readlength)
```

bone studies
```{r}
bone_n_rse_adj <- bone_n_rse %>% 
  mutate(., source = "BONE", origin = "NonDisease Tissue", rse_samp_name = rownames(bone_n_rse), tissue = "Bone", cell_line = NA, disease = "NonDisease Bone", 
         sequencer = "Illumina HiSeq", sex = "Not Reported", race = "Not Reported", readlength = `recount_qc.star.average_input_read_length`, .keep = "used") %>% 
  dplyr::select(source, origin, rse_samp_name, tissue, cell_line, disease, sequencer, sex, race, readlength)

bone_c_rse_adj <- bone_c_rse %>% 
  mutate(., source = "BONE", origin = "Cell Line", rse_samp_name = rownames(bone_c_rse), tissue = "Bone", cell_line = barebones_cellline(`sra_attribute.source_name`), 
         disease = "Osteosarcoma", sequencer = "Illumina HiSeq", sex = "Not Reported", race = "Not Reported", readlength = `recount_qc.star.average_input_read_length`, 
         .keep = "used") %>%
  dplyr::select(source, origin, rse_samp_name, tissue, cell_line, disease, sequencer, sex, race, readlength)

bone_t_rse_adj <- bone_t_rse %>% 
  mutate(., source = "BONE", origin = "Tumor Tissue", rse_samp_name = rownames(bone_t_rse), tissue = "Bone", cell_line = NA, disease = "Osteosarcoma", 
         sequencer = "Illumina HiSeq", sex = "Not Reported", race = "Not Reported", readlength = `recount_qc.star.average_input_read_length`, .keep = "used") %>% 
  dplyr::select(source, origin, rse_samp_name, tissue, cell_line, disease, sequencer, sex, race, readlength)
```

human protein atlas
```{r}
hpa_c_rse_adj <- hpa_c_rse %>% 
  mutate(., source = "HPA CL", origin = "Cell Line", rse_samp_name = rownames(hpa_c_rse), tissue = hpa_tissues, cell_line = str_sub(`sra.sample_name`, end = -2), 
         disease = str_to_title(str_sub(`sra.sample_description`, end = -10)), sequencer = "Illumina HiSeq", sex = "Not Reported", race = "Not Reported", 
         readlength = `recount_qc.star.average_input_read_length`, .keep = "used") %>% 
  dplyr::select(source, origin, rse_samp_name, tissue, cell_line, disease, sequencer, sex, race, readlength)

hpa_t_rse_adj <- hpa_t_rse %>% 
  mutate(., source = "HPA T", origin = "NonDisease Tissue", rse_samp_name = rownames(hpa_t_rse), tissue = str_to_title(`sra_attribute.organism_part`), cell_line = NA, 
         disease = paste("NonDisease", str_to_title(`sra_attribute.organism_part`)), sequencer = "Illumina HiSeq", sex = "Not Reported", race = "Not Reported", 
         readlength = `recount_qc.star.average_input_read_length`, .keep = "used") %>%
  dplyr::select(source, origin, rse_samp_name, tissue, cell_line, disease, sequencer, sex, race, readlength)
hpa_t_rse_adj[hpa_t_rse_adj$tissue == "Homo Sapiens",]$tissue <- str_to_title(hpa_t_rse[hpa_t_rse$sra_attribute.organism_part == "Homo sapiens",]$sra_attribute.INSDC_status) # a few cells' columns of this RSE are shifted to the left for some reason

```

CCLE and PDX
```{r}
ccle_rse_adj <- ccle_rse %>% 
  mutate(., source ="CCLE", origin = "Cell Line", rse_samp_name = rownames(ccle_rse), tissue = str_to_title(gsub("_", " ", `sra_attribute.tissue`)), 
         cell_line = barebones_cellline(sra_attribute.cell_line), disease = str_to_title(gsub("_", " ", `sra_attribute.disease`)), sequencer = "Illumina HiSeq", 
         sex = ifelse(sra_attribute.sex == "NA", "Not Reported", str_to_title(sra_attribute.sex)), race = ifelse(sra_attribute.ethnicity == "NA", "Not Reported", `sra_attribute.ethnicity`), readlength = `recount_qc.star.average_input_read_length`, .keep = "used") %>% 
  #replace_na(list(sex = "Not Reported", race = "Not Reported")) %>%
  dplyr::select(source, origin, rse_samp_name, tissue, cell_line, disease, sequencer, sex, race, readlength)

pdx_rse_adj <- pdx_rse %>% 
  mutate(., source = "PDX", origin = "PDX", rse_samp_name = rownames(pdx_rse), tissue = "Brain", cell_line = NA, disease = "Glioblastoma", sequencer = "Illumina HiSeq", 
         sex = str_to_title(sra_attribute.sex), race = "Not Reported", readlength = `recount_qc.star.average_input_read_length`, .keep = "used") %>% 
  dplyr::select(source, origin, rse_samp_name, tissue, cell_line, disease, sequencer, sex, race, readlength)

```

combine and adjust
```{r}
sample_master_meta <- rbind(tcga_meta_adj, gtex_meta_adj, bone_c_rse_adj, bone_n_rse_adj, bone_t_rse_adj, hpa_c_rse_adj, hpa_t_rse_adj, ccle_rse_adj, pdx_rse_adj)
saveRDS(sample_master_meta, "../data/combined_projects_metadata.rds")

# adjust some wording in the tissue_coerced and disease_coerced columns
sample_master_meta$tissue_coerced <- stri_replace_all_regex(sample_master_meta$tissue,
                                             pattern = c("Haematopoietic And Lymphoid Tissue|Lymph Nodes|Bone Marrow|Lymph Node", "Bile Duct|Biliary Tract", "Autonomic Ganglia", 
                                                         "Upper Aerodigestive Tract", "Central Nervous System|Cerebral Cortex", "Large Intestine|Colon Adenocarcinoma", 
                                                         "Urinary Tract", "Endometrium", "Oesophagus", "Cervix Uteri", "Rectum Adenocarcinoma", "Gall Bladder", "Duodenum", 
                                                         "Animal Ovary"),
                                             replacement = c("Blood", "Biliary System", "Nerve", "Head and Neck", "Brain", "Colon", "Bladder", "Uterus", "Esophagus", "Cervix", 
                                                             "Rectum", "Gallbladder", "Small Intestine", "Ovary"),
                                             vectorize = FALSE)

sample_master_meta <- sample_master_meta %>% mutate(tissue_coerced = ifelse(tissue %in% "Colorectal", ifelse(disease %in% "Colon Adenocarcinoma", "Colon", ifelse(disease %in% "Rectum Adenocarcinoma", "Rectum", tissue_coerced)), tissue_coerced))

## now add a column that coerces diseases into broader groups
sample_master_meta <- sample_master_meta %>% mutate(disease_coerced = ifelse(origin %in% c("Cell Line", "PDX", "Tumor Tissue"), paste(tissue_coerced, "Cancer"), paste("NonDisease", tissue_coerced)))

#grammar corrections 
sample_master_meta <- sample_master_meta %>% mutate(disease_coerced = ifelse(disease_coerced %in% "Uterus Cancer", "Uterine Cancer",
                                                                             ifelse(disease_coerced %in% "Ovary Cancer", "Ovarian Cancer",
                                                                                    ifelse(disease_coerced %in% "Esophagus Cancer", "Esophageal Cancer",
                                                                                           ifelse(disease_coerced %in% "Adrenal Gland Cancer", "Adrenal Cancer",
                                                                                                  ifelse(disease_coerced %in% "Pancreas Cancer", "Pancreatic Cancer",
                                                                                                         ifelse(disease_coerced %in% "Rectum Cancer", "Rectal Cancer",
ifelse(disease_coerced %in% "Testis Cancer", "Testicular Cancer",
       ifelse(disease_coerced %in% "Pleura Cancer", "Pleural Cancer",
              ifelse(disease_coerced %in% "Cervix Cancer", "Cervical Cancer",
                     ifelse(disease_coerced %in% "Biliary System Cancer", "Biliary Cancer", disease_coerced)))))))))))

```

#save
```{r}
sample_master_meta <- rownames_to_column(sample_master_meta, var = "rse_sampname")
saveRDS(sample_master_meta, "../data/sample_master_meta.rds")
```

#compare results (remove chunk after code review)
```{r}
averys_meta <- readRDS("../data/OLD_sample_master_meta.rds")
##compare
BiocManager::install("arsenal") #not in docker image , just for checking results
library(arsenal) 
averys_meta$tissue_groups <- NULL #removed, not used
sample_master_meta$tissue_groups <- NULL 


meta_differences <- comparedf(sample_master_meta, averys_meta, by = "rse_sampname") 
summary(meta_differences)
View(meta_differences)
View(meta_differences[["vars.summary"]][["values"]][[12]]) #check differences by columns
#tolerable differences in 3, 4, 5, 6, 8, 9, 11, 12:
#after updates to metadata, found differences compared to the original in 'control' -> 'nondisease' rewording, differences in tissue (but not carried over to tissue_coerced, which is the only thing used), differences in cell line upper/lower case, differences in tissue upper/lower case
```

