---
title: "SubsetRSE"
output: html_document
date: '2022-08-10'
---

> script to generate subset RSEs from a full RSE based on nuumber of most variable genes to include, whether variation is based on TCGA or GTEx, and whether to exclude purity genes

# Prerequisites: objects with GTEx and TCGA TPM and all RSE objects, which are exported in the DataDownload script, rse_to_subset.R function script
```{r source load}
source("../scripts/functions/rse_to_subset.R") # change to where you have this function script saved
```

# make objects with varying genes for TCGA and GTEx to use with function
## computation for this sourced from https://github.com/katharineyu/TCGA_CCLE_paper/blob/master/R_scripts/tumor_cell_line_cor.R
```{r var genes}
iqr_gene <- apply(tcga_tpm, 1, IQR, na.rm = TRUE)
tcga_varying_genes <- iqr_gene[order(iqr_gene, decreasing = T)]

iqr_gene <- apply(gtex_tpm, 1, IQR, na.rm = TRUE)
gtex_varying_genes <- iqr_gene[order(iqr_gene, decreasing = T)]

purity_genes <- readRDS("../data/purity_genes.rds") # replace filepath
```

# use function in rse_to_subset.R script to create RSE subsets accounting for each variable (RSE source, subset size, in- or excluding tumor purity-correlated genes, basing variant gene rankings on TCGA vs GTEx)
## this is set up to save each object but not to place in working env
```{r run function}
project <- c("tcga", "gtex", "ccle", "hpa_t", "hpa_c", "pdx", "bone_t", "bone_n", "bone_c")
subset_num <- c(100, 1000, 5000, 10000)
var_source <- c("tcga", "gtex")

for(i in project){
  project_sub <- i
  rse_sub <- readRDS(paste("../data/recount3/rse", "/", project_sub, "_rse.rds", sep = ""))
  
  for(i in subset_num){
    subset_sub <- i
    
    for(i in var_source){
      var_sub <- i
      
      rse_to_subset(rse_sub, project_sub, subset_sub, var_sub, TRUE, TRUE, TRUE, "../data/rse_subsets/") # with purity
      rse_to_subset(rse_sub, project_sub, subset_sub, var_sub, FALSE, TRUE, TRUE, "../data/rse_subsets/") # without
    }
  }
}
```

# clean environment
```{r clean environment}
rm(iqr_gene, tcga_varying_genes, gtex_varying_genes, project, subset_num, var_source, rse_sub, subset_sub, var_sub, project_sub, i, purity_genes)
```