# incorporating GTEx tissue-specific genes (https://maayanlab.cloud/Harmonizome/dataset/GTEx+Tissue+Gene+Expression+Profiles)

# requires qusage package for reading in GMTs
```{r read in up and down genes for each tissue}
gtex_up_tissue_genes <- qusage::read.gmt("../data/gene_set_library_up_crisp.gmt")
gtex_down_tissue_genes <- qusage::read.gmt("../data/gene_set_library_dn_crisp.gmt")
sample_master_meta <- readRDS("../data/sample_master_meta.rds")
```

# I need ens IDs for these genes
```{r adding ens IDs}
library(AnnotationDbi)
library(org.Hs.eg.db)

for(i in names(gtex_up_tissue_genes)){
  out <- data.frame(genesym = gtex_up_tissue_genes[[paste(i)]])
  out$ens <- mapIds(org.Hs.eg.db, keys = out$genesym, column = "ENSEMBL", keytype = "SYMBOL", multiVals = "first")
 gtex_up_tissue_genes[[paste(i)]] <- na.omit(out$ens) 
}

for(i in names(gtex_down_tissue_genes)){
  out <- data.frame(genesym = gtex_down_tissue_genes[[paste(i)]])
  out$ens <- mapIds(org.Hs.eg.db, keys = out$genesym, column = "ENSEMBL", keytype = "SYMBOL", multiVals = "first")
 gtex_down_tissue_genes[[paste(i)]] <- na.omit(out$ens) 
}

# also would like to have one list with both up and down
gtex_both_tissue_genes <- list()
for(i in names(gtex_up_tissue_genes)){ # up and down have the same name set
  gtex_both_tissue_genes[[paste(i)]] <- c(gtex_up_tissue_genes[[paste(i)]], gtex_down_tissue_genes[[paste(i)]])
}

# clear env
rm(out, i)
rm(gtex_down_tissue_genes, gtex_up_tissue_genes) # optional
```
# next, subset this nested list to only the cell line tissue origins represented
```{r list subset}
cellline_tissues <- unique(sample_master_meta[sample_master_meta$origin == "Cell Line",]$tissue_coerced)
gtex_tissues_cap <- stringr::str_to_title(names(gtex_both_tissue_genes))

# what tissues are in common
#View(data.frame(intersect(cellline_tissues, gtex_tissues_cap))) # 20 in common so far
#View(data.frame(setdiff(cellline_tissues, gtex_tissues_cap))) # Head And Neck, Pleura, Bone, Soft Tissue, Biliary System in cl but not gtex
#View(data.frame(setdiff(gtex_tissues_cap, cellline_tissues))) # Heart, Muscle, Adipose, Blood Vessel, Fallopian, Vagina, Pituitary, Adrenal, Spleen in gtex but not cl

# making list of tissues to exclude from gtex
gtex_tissues_exc <- stringr::str_to_lower(setdiff(gtex_tissues_cap, cellline_tissues)) 

# remove these from my list
for(i in gtex_tissues_exc){
  gtex_both_tissue_genes[[paste(i)]] <- NULL
}

# optional check
gtex_tissues_fin <- "x"
for(i in 1:length(gtex_both_tissue_genes)){
  gtex_tissues_fin <- c(gtex_tissues_fin, names(gtex_both_tissue_genes[i]))
}
gtex_tissues_fin <- gtex_tissues_fin[-1]
#View(data.frame(intersect(cellline_tissues, stringr::str_to_title(gtex_tissues_fin)))) # all but intestine

# remove objects
rm(gtex_tissues_cap, gtex_tissues_exc, cellline_tissues, i)
```

# now, use function to correlate x by y with method including gene set
```{r tissue gene set cor fun}
source("../scripts/functions/tisgenesetcor.R")
```

# make tpm objs to run fun
## pull in tpm matrices
```{r pull tpm}
tpm_to_pull <- c("gtex", "tcga", "ccle", "bone_t", "hpa_t", "pdx", "hpa_c", "bone_c", "bone_n")

for(i in tpm_to_pull){
  assign(paste(i, "_tpm", sep = ""), readRDS(paste("../data/recount3/tpm/", i, "_tpm.rds", sep = "")))
}

rm(tpm_to_pull, i)
```

## put together
```{r bind tpm matrices}
# match order with meta table:
## PM: CCLE, gtex Cell Lines, PDX, BONE CL
# T: TCGA, GTEx, gtex Tissues, BONE Tumor, BONE Normal

cl_tpm <- cbind(ccle_tpm, hpa_c_tpm, bone_c_tpm) # NOTE: JUST including cell lines

# need to subset cl_tpm to only samples with tissues represented in GTEx
cl_tpm <- cl_tpm[,colnames(cl_tpm) %in% sample_master_meta[sample_master_meta$tissue_coerced %in% stringr::str_to_title(gtex_tissues_fin),]$rse_sampname]

con_tissue_tpm <- cbind(gtex_tpm, hpa_t_tpm, bone_n_tpm)
tum_tissue_tpm <- cbind(tcga_tpm, bone_t_tpm)
```

# run function
```{r run fun}
corbytis_tum_res <- list()
for(i in gtex_tissues_fin){
  corbytis_tum_res[[paste(i)]] <- tisgenesetcor(cl_tpm, tum_tissue_tpm, tissue = i)
}

corbytis_con_res <- list()
for(i in gtex_tissues_fin){
  corbytis_con_res[[paste(i)]] <- tisgenesetcor(cl_tpm, con_tissue_tpm, tissue = i)
}

rm(i)
```

# visualize as violin plot
## first, need to make the data length-wise
```{r prep data for violin plot}
tum_corres <- 1
for(i in gtex_tissues_fin){
  tum_corres <- c(tum_corres, corbytis_tum_res[[paste(i)]]$cor)
}
tum_corres <- tum_corres[-1]

con_corres <- 1
for(i in gtex_tissues_fin){
  con_corres <- c(con_corres, corbytis_con_res[[paste(i)]]$cor)
}
con_corres <- con_corres[-1]

tissuerep <- "x"
for(i in gtex_tissues_fin){
  tissuerep <- c(tissuerep, rep(i, length(corbytis_tum_res[[paste(i)]]$cor))) # tissue will be the same for tum/con so don't need split
}
tissuerep <- tissuerep[-1]

tis_gene_cor_tum_vio_in <- data.frame(corres = tum_corres, tissue = stringr::str_to_title(tissuerep))
# since salivary gland is blank, let's just remove
tis_gene_cor_tum_vio_in <- tis_gene_cor_tum_vio_in[tis_gene_cor_tum_vio_in$tissue != "Salivary Gland",]
tis_gene_cor_tum_vio_in <- tis_gene_cor_tum_vio_in[tis_gene_cor_tum_vio_in$tissue != "Small Intestine",]
tis_gene_cor_tum_vio_in <- tis_gene_cor_tum_vio_in[tis_gene_cor_tum_vio_in$tissue != "Nerve",] # same for nerve

tis_gene_cor_con_vio_in <- data.frame(corres = con_corres, tissue = stringr::str_to_title(tissuerep))
#tis_gene_cor_con_vio_in <- tis_gene_cor_con_vio_in[tis_gene_cor_con_vio_in$tissue != "Salivary Gland",]
tis_gene_cor_con_vio_in <- tis_gene_cor_con_vio_in[tis_gene_cor_con_vio_in$tissue != "Small Intestine",]
```

## next, generate plots
```{r generate violin plot}
library(ggplot2)
ggplot(tis_gene_cor_tum_vio_in, aes(x=tissue, y=corres, color=tissue)) + 
  geom_violin() + geom_jitter(shape=16, position=position_jitter(0.2)) + labs(x = "Matched Tumor Tissue", y = "Median Correlation of Each Cell Line Sample to Matched Tumor Tissue") + ylim(0.4,0.9)
ggsave(filename = "../plots/tags_violin/allcltomatchtum.png", 
          width = 13, height = 7, units = "in")

ggplot(tis_gene_cor_con_vio_in, aes(x=tissue, y=corres, color=tissue)) + 
  geom_violin() + geom_jitter(shape=16, position=position_jitter(0.2)) + labs(x = "Matched Non-Disease Tissue", y = "Median Correlation of Each Cell Line Sample to Matched Control Tissue") + ylim(0.4,0.9)
ggsave(filename = "../plots/tags_violin/allcltomatchcon.png", 
          width = 13, height = 7, units = "in")
```

# split violin
## geom_split_violin fun
```{r}
source("../scripts/functions/geom_split_violin.R")
```

# plot split
```{r}
# make combined obj with tum and con
tis_gene_cor_both_vio_in <- rbind(tis_gene_cor_tum_vio_in, tis_gene_cor_con_vio_in)
tis_gene_cor_both_vio_in$Type <- c(rep("Tumor", length(tis_gene_cor_tum_vio_in$corres)), rep("Non-Disease", length(tis_gene_cor_con_vio_in$corres)))
tis_gene_cor_both_vio_in <- tis_gene_cor_both_vio_in[tis_gene_cor_both_vio_in$tissue != "Nerve",]
tis_gene_cor_both_vio_in <- tis_gene_cor_both_vio_in[tis_gene_cor_both_vio_in$tissue != "Salivary Gland",]

ggplot(tis_gene_cor_both_vio_in, aes(tissue, corres, fill = Type)) + geom_split_violin(width = 1) + labs(x = "Tissue of Origin", y = "Median Correlation Between Matched Cell Lines and Tissues") + theme(axis.text.x = element_text(angle = 45, hjust=1)) + scale_fill_manual(values = c("darkblue", "orange")) + labs(fill = "Tissue Type") + theme(text = element_text(face = "bold"), legend.text = element_text(size = 16), legend.title = element_text(size = 18), axis.title = element_text(size = 12.5), axis.text = element_text(size = 15))
ggsave(filename = "../plots/tags_violin/allcltomatchboth.png", 
          width = 13, height = 7, units = "in")
```
#### Wilcoxon rank sum test on some of these groups
```{r}
# Question : Is there any significant difference between pdx and cell line median correlation to matched tissue when only considering TAGs?
library(tidyverse)
library(rstatix)
library(ggpubr)
library(coin)

# make input object
tis_gene_cor_stat <- list()

for(i in unique(tis_gene_cor_tum_vio_in$tissue)){# I know the same tissues are represented in tum vs con
  tis_gene_cor_stat[[i]][["corres"]] <- data.frame(med_cor = c(tis_gene_cor_tum_vio_in[tis_gene_cor_tum_vio_in$tissue == i,]$corres, tis_gene_cor_con_vio_in[tis_gene_cor_con_vio_in$tissue == i,]$corres), type = c(rep("Tumor", length(tis_gene_cor_tum_vio_in[tis_gene_cor_tum_vio_in$tissue == i,]$corres)), rep("Control", length(tis_gene_cor_con_vio_in[tis_gene_cor_con_vio_in$tissue == i,]$corres))))
}

# compute effect size for each
for(i in names(tis_gene_cor_stat)){
  # computation
  tis_gene_cor_stat[[i]][["effect"]] <- tis_gene_cor_stat[[i]][["corres"]] %>%
    rstatix::wilcox_test(med_cor ~ type) %>%
    add_significance()
  # effect size
  tis_gene_cor_stat[[i]][["corres"]] %>% wilcox_effsize(med_cor ~ type)
}

# view in dataframe form
temp <- data.frame(tissue = names(tis_gene_cor_stat), effect = NA)
for(i in temp$tissue){
  temp[grep(i, temp$tissue),]$effect <- tis_gene_cor_stat[[i]][["effect"]][["p"]]
}
View(temp)
rm(temp)
```

## next, let's make a plot for each tissue
### how did each cell line recapitulate another tissue?
```{r}
corbytis_tum_allcl_res <- list()
for(i in gtex_tissues_fin){
  corbytis_tum_allcl_res[[paste(i)]] <- tisgenesetcor(cl_tpm, tum_tissue_tpm, tissue = i, onlymatch = FALSE)
}

corbytis_con_allcl_res <- list()
for(i in gtex_tissues_fin){
  corbytis_con_allcl_res[[paste(i)]] <- tisgenesetcor(cl_tpm, con_tissue_tpm, tissue = i, onlymatch = FALSE)
}
```

## now, make new objects that split each nested list by cell line origin tissue (nested 1 = tissue/TAGs of comparison, nested 2 = model tissue of origin)
```{r}
for(i in 1:length(corbytis_tum_allcl_res)){
  tissue <- gtex_tissues_fin[i]
  temp <- corbytis_tum_allcl_res[[paste(tissue)]]
  samples <- rownames(temp)
  cl_split <- list()
  master_sub <- sample_master_meta[sample_master_meta$rse_sampname %in% samples,]
  for(i in gtex_tissues_fin){
    if(i != "intestine"){
      samples_sub <- master_sub[master_sub$tissue_coerced == stringr::str_to_title(i),]$rse_sampname
    } else if(i == "intestine"){
      samples_sub <- c(master_sub[master_sub$tissue_coerced == "Colon",]$rse_sampname, master_sub[master_sub$tissue_coerced == "Small Intestine",]$rse_sampname)
    }
    cl_split[[paste(i)]] <- temp[rownames(temp) %in% samples_sub,]
  }
  corbytis_tum_allcl_res[[paste(tissue)]] <- cl_split
  rm(temp,tissue,samples,cl_split,master_sub,samples_sub)
}

for(i in 1:length(corbytis_con_allcl_res)){
  tissue <- gtex_tissues_fin[i]
  temp <- corbytis_con_allcl_res[[paste(tissue)]]
  samples <- rownames(temp)
  cl_split <- list()
  master_sub <- sample_master_meta[sample_master_meta$rse_sampname %in% samples,]
  for(i in gtex_tissues_fin){
    if(i != "intestine"){
      samples_sub <- master_sub[master_sub$tissue_coerced == stringr::str_to_title(i),]$rse_sampname
    } else if(i == "intestine"){
      samples_sub <- c(master_sub[master_sub$tissue_coerced == "Colon",]$rse_sampname, master_sub[master_sub$tissue_coerced == "Small Intestine",]$rse_sampname)
    }
    cl_split[[paste(i)]] <- temp[rownames(temp) %in% samples_sub,]
  }
  corbytis_con_allcl_res[[paste(tissue)]] <- cl_split
  rm(temp,tissue,samples,cl_split,master_sub,samples_sub)
}
```

## make objects for violin plot
```{r}
allcl_totistags_vio_in <- list()
for(i in gtex_tissues_fin){
  tum_allcl_corres <- 1
  tissue <- i
  for(i in gtex_tissues_fin){
    tum_allcl_corres <- c(tum_allcl_corres, corbytis_tum_allcl_res[[paste(tissue)]][[paste(i)]])
  }
  tum_allcl_corres <- tum_allcl_corres[-1]
  
  con_allcl_corres <- 1
  for(i in gtex_tissues_fin){
    con_allcl_corres <- c(con_allcl_corres, corbytis_con_allcl_res[[paste(tissue)]][[paste(i)]])
  }
  con_allcl_corres <- con_allcl_corres[-1]
  
  tissue_allcl_rep <- "x"
  for(i in gtex_tissues_fin){
    tissue_allcl_rep <- c(tissue_allcl_rep, rep(i, length(corbytis_tum_allcl_res[[paste(tissue)]][[paste(i)]]))) # tissue will be the same for tum/con so don't need split
  }
  tissue_allcl_rep <- tissue_allcl_rep[-1]
  
  allcl_totistags_vio_in[[paste(tissue)]][["tum"]] <- data.frame(corres = tum_allcl_corres, tissue = tissue_allcl_rep)
  allcl_totistags_vio_in[[paste(tissue)]][["con"]] <- data.frame(corres = con_allcl_corres, tissue = tissue_allcl_rep)
  
  # clean env
  rm(tum_allcl_corres, tissue, con_allcl_corres, tissue_allcl_rep)
}
```

## make and export a buncha violin plots
```{r}
library(ggplot2)
violin_outputs <- list()
for(i in c("tum", "con")){
  type <- i
  for(i in gtex_tissues_fin){
    if(type == "tum"){
      ggplot(allcl_totistags_vio_in[[paste(i)]][[paste(type)]], aes(x=tissue, y=corres, color=tissue)) + 
  geom_violin() + geom_jitter(shape=16, position=position_jitter(0.2)) + ggtitle(paste("Median Correlation of Cell Lines of Each Tissue Type to ", stringr::str_to_title(i), " Tumor Tissue by ", stringr::str_to_title(i), " TAGs", sep = ""))
      ggsave(filename = paste("../plots/tags_violin/", i, "tum.png", sep = ""), 
          width = 15, height = 7, units = "in")
    } else if(type == "con"){
      ggplot(allcl_totistags_vio_in[[paste(i)]][[paste(type)]], aes(x=tissue, y=corres, color=tissue)) + 
  geom_violin() + geom_jitter(shape=16, position=position_jitter(0.2)) + ggtitle(paste("Median Correlation of Cell Lines of Each Tissue Type to ", stringr::str_to_title(i), " Control Tissue by ", stringr::str_to_title(i), " TAGs", sep = ""))
      ggsave(filename = paste("../plots/tags_violin/", i, "con.png", sep = ""), 
          width = 15, height = 7, units = "in")
    }
  }
}
```

# add PDX vs brain cl performance
## first, how they compare in brain tissue
```{r}
# make TPM obj for just GBM CLs
GBM_cl_tpm <- cl_tpm[,colnames(cl_tpm) %in% sample_master_meta[grep("Grade Iv|Glioblastoma", sample_master_meta$disease),]$rse_sampname]

# run function
# computes performance of GBM cls vs GBM pdxs in correlating to brain tissue in brain TAGs
corbytis_pdxcl_tum_res <- tisgenesetcor(cbind(GBM_cl_tpm, pdx_tpm), tum_tissue_tpm, tissue = "brain")

corbytis_pdxcl_con_res <- tisgenesetcor(cbind(GBM_cl_tpm, pdx_tpm), con_tissue_tpm, tissue = "brain")
```

### visualize as violin plot
#### first, need to make the data length-wise
```{r prep data for violin plot}
originrep <- c(rep("CL", ncol(GBM_cl_tpm)), rep("PDX", ncol(pdx_tpm)))

tis_gene_cor_pdxcl_tum_vio_in <- data.frame(corres = corbytis_pdxcl_tum_res$cor, type = originrep)
tis_gene_cor_pdxcl_tum_vio_in$type[tis_gene_cor_pdxcl_tum_vio_in$type == "CL"] <- "Cell Line"

tis_gene_cor_pdxcl_con_vio_in <- data.frame(corres = corbytis_pdxcl_con_res$cor, type = originrep)
tis_gene_cor_pdxcl_con_vio_in$type[tis_gene_cor_pdxcl_con_vio_in$type == "CL"] <- "Cell Line"
```

#### next, generate plots
```{r generate violin plot}
library(ggplot2)
ggplot(tis_gene_cor_pdxcl_tum_vio_in, aes(x=type, y=corres, color=type)) + geom_violin() + geom_jitter(shape=16, position=position_jitter(0.2)) + labs(x = "Model Type", y = "Median Correlation of Model to Brain Tumor Tissue in Brain-Specific Genes") + ylim(0.5,0.9) + guides(color = "none") + theme(text = element_text(face = "bold"), legend.text = element_text(size = 16), legend.title = element_text(size = 18), axis.title = element_text(size = 12), axis.text = element_text(size = 15))
ggsave(filename = "../plots/tags_violin/brainclpdxtomatchtum.png", 
          width = 6, height = 7)

print("PDX median correlation to tumor tissue rho is")
median(tis_gene_cor_pdxcl_tum_vio_in[tis_gene_cor_pdxcl_tum_vio_in$type == "PDX",]$corres) - median(tis_gene_cor_pdxcl_tum_vio_in[tis_gene_cor_pdxcl_tum_vio_in$type == "Cell Line",]$corres)
print("greater than Cell Line median")

ggplot(tis_gene_cor_pdxcl_con_vio_in, aes(x=type, y=corres, color=type)) + geom_violin() + geom_jitter(shape=16, position=position_jitter(0.2)) + labs(x = "Model Type", y = "Median Correlation of Model to Brain Non-Disease Tissue in Brain-Specific Genes") + ylim(0.5,0.9) + guides(color = "none") + theme(text = element_text(face = "bold"), legend.text = element_text(size = 16), legend.title = element_text(size = 18), axis.title = element_text(size = 12), axis.text = element_text(size = 15))
ggsave(filename = "../plots/tags_violin/brainclpdxtomatchcon.png", 
          width = 6, height = 7)

print("PDX median correlation to non-disease tissue rho is")
median(tis_gene_cor_pdxcl_con_vio_in[tis_gene_cor_pdxcl_con_vio_in$type == "PDX",]$corres) - median(tis_gene_cor_pdxcl_con_vio_in[tis_gene_cor_pdxcl_con_vio_in$type == "Cell Line",]$corres)
print("greater than Cell Line median")
```

#### Wilcoxon rank sum test on these two groups
```{r}
# Question : Is there any significant difference between pdx and cell line median correlation to matched tissue when only considering TAGs?
library(tidyverse)
library(rstatix)
library(ggpubr)
library(coin)

# make input object
tis_gene_cor_pdxcl_stat <- data.frame(tum_med_cor = tis_gene_cor_pdxcl_tum_vio_in$corres, con_med_cor = tis_gene_cor_pdxcl_con_vio_in$corres, type = tis_gene_cor_pdxcl_tum_vio_in$type)

# computation - tumor
stat.test_tag <- tis_gene_cor_pdxcl_stat %>% 
  rstatix::wilcox_test(tum_med_cor ~ type) %>%
  add_significance()
stat.test_tag

# effect size
tis_gene_cor_pdxcl_stat %>% wilcox_effsize(tum_med_cor ~ type)

# report
stat.test_tag <- stat.test_tag %>% add_xy_position(x = "group")
tis_gene_cor_pdxcl_tum_v_plot + 
  stat_pvalue_manual(stat.test_tag, tip.length = 0) +
  labs(subtitle = get_test_label(stat.test_tag, detailed = TRUE))
#ggsave(filename = "../plots/tags_violin/brainclpdxtomatchtum.png", 
         # width = 5, height = 6)

# computation - control
stat.test_tag <- tis_gene_cor_pdxcl_stat %>% 
  rstatix::wilcox_test(con_med_cor ~ type) %>%
  add_significance()
stat.test_tag

# effect size
library(coin)
tis_gene_cor_pdxcl_stat %>% wilcox_effsize(con_med_cor ~ type)

# report
library(ggpubr)
stat.test_tag <- stat.test_tag %>% add_xy_position(x = "group")
tis_gene_cor_pdxcl_con_v_plot + 
  stat_pvalue_manual(stat.test_tag, tip.length = 0) +
  labs(subtitle = get_test_label(stat.test_tag, detailed = TRUE))
#ggsave(filename = "../plots/tags_violin/brainclpdxtomatchcon.png", 
         # width = 5, height = 6)
```

## next, how they compare in recapitulating expression across tissue TAGs
### first, correlation of each across each tissue
```{r}
corbytis_pdxcl_alltum_res <- list()
corbytis_pdxcl_allcon_res <- list()

for(i in c("tum", "con")){
  type <- i
  
  for(i in c("cl", "pdx")){
    model <- i
    
    if(i == "cl"){
      x <- GBM_cl_tpm
    }else if(i == "pdx"){
      x <- pdx_tpm
    }
    
    for(i in gtex_tissues_fin){
      tissue <- i
      if(type == "tum"){
        corbytis_pdxcl_alltum_res[[paste(tissue)]][[paste(model)]] <- tisgenesetcor(x, tum_tissue_tpm, tissue = tissue, onlymatch = FALSE)
      } else if(type == "con"){
        corbytis_pdxcl_allcon_res[[paste(tissue)]][[paste(model)]] <- tisgenesetcor(x, con_tissue_tpm, tissue = tissue, onlymatch = FALSE)
      }
    }
  }
}
```

### next, build data frames for inputs to ggplot
```{r}
tis_gene_cor_pdxcl_allcon_vio_in <- data.frame(cor = 1, type = "x", tissue = "x")
tis_gene_cor_pdxcl_alltum_vio_in <- data.frame(cor = 1, type = "x", tissue = "x")
for(i in c("tum", "con")){
  tis <- i
  
  for(i in gtex_tissues_fin){
    tissue <- i
    if(tis == "tum"){
      sub <- corbytis_pdxcl_alltum_res[[paste(tissue)]]
    } else if(tis == "con"){
      sub <- corbytis_pdxcl_allcon_res[[paste(tissue)]]
    }
    
    cor <- c(sub[["cl"]]$cor, sub[["pdx"]]$cor)
    type <- c(rep("Cell Line", length(sub[["cl"]]$cor)), rep("PDX", length(sub[["pdx"]]$cor)))
    tissue <- rep(paste(tissue), length(cor))
    
    if(tis == "tum"){
      tis_gene_cor_pdxcl_alltum_vio_in <- rbind(tis_gene_cor_pdxcl_alltum_vio_in, data.frame(cor = cor, type = type, tissue = stringr::str_to_title(tissue)))
    } else if(tis == "con"){
      tis_gene_cor_pdxcl_allcon_vio_in <- rbind(tis_gene_cor_pdxcl_allcon_vio_in, data.frame(cor = cor, type = type, tissue = stringr::str_to_title(tissue)))
    }
  }
}
tis_gene_cor_pdxcl_allcon_vio_in <- tis_gene_cor_pdxcl_allcon_vio_in[-1,]
tis_gene_cor_pdxcl_alltum_vio_in <- tis_gene_cor_pdxcl_alltum_vio_in[-1,]

# also remove salivary gland from tumor
tis_gene_cor_pdxcl_alltum_vio_in <- tis_gene_cor_pdxcl_alltum_vio_in[tis_gene_cor_pdxcl_alltum_vio_in$tissue != "Salivary Gland",]
tis_gene_cor_pdxcl_alltum_vio_in <- tis_gene_cor_pdxcl_alltum_vio_in[tis_gene_cor_pdxcl_alltum_vio_in$tissue != "Nerve",]
tis_gene_cor_pdxcl_alltum_vio_in <- tis_gene_cor_pdxcl_alltum_vio_in[tis_gene_cor_pdxcl_alltum_vio_in$tissue != "Small Intestine",]
```

### make plots

#### run fun to make objects
```{r}
library(ggplot2)
ggplot(tis_gene_cor_pdxcl_alltum_vio_in, aes(tissue, cor, fill = type)) + geom_split_violin() + labs(x = "Tumor Tissue Type", y = "Median Correlation of GBM Models to Tissues in Each Tissue's Specific Genes") + theme(axis.text.x = element_text(angle = 45, hjust=1)) + scale_fill_discrete(name = "Model") + theme(text = element_text(face = "bold"), legend.text = element_text(size = 16), legend.title = element_text(size = 18), axis.title = element_text(size = 12.5), axis.text = element_text(size = 15))
ggsave(filename = "../plots/tags_violin/brainclpdxtoalltum.png", 
          width = 5000, height = 2300, units = "px")

range(tis_gene_cor_pdxcl_alltum_vio_in[tis_gene_cor_pdxcl_alltum_vio_in$type == "Cell Line",]$cor)
range(tis_gene_cor_pdxcl_alltum_vio_in[tis_gene_cor_pdxcl_alltum_vio_in$type == "PDX",]$cor)

ggplot(tis_gene_cor_pdxcl_allcon_vio_in, aes(tissue, cor, fill = type)) + geom_split_violin() + labs(x = "Control Tissue Type", y = "Median Correlation of GBM Models to Tissues in Each Tissue's Specific Genes") + theme(axis.text.x = element_text(angle = 45, hjust=1)) + scale_fill_discrete(name = "Model") + theme(text = element_text(face = "bold"), legend.text = element_text(size = 16), legend.title = element_text(size = 18), axis.title = element_text(size = 12.5), axis.text = element_text(size = 15))
ggsave(filename = "../plots/tags_violin/brainclpdxtoallcon.png", 
          width = 5000, height = 2300, units = "px")
```

# make heatmap to summarize all cls compared to each tissue/TAG
```{r}
corbytis_tum_allcl_df <- tisgenesetcor(cl_tpm, tum_tissue_tpm, tissue = gtex_tissues_fin[1], onlymatch = FALSE)
for(i in gtex_tissues_fin[-1]){
  corbytis_tum_allcl_df <- cbind(corbytis_tum_allcl_df, tisgenesetcor(cl_tpm, tum_tissue_tpm, tissue = i, onlymatch = FALSE))
}
colnames(corbytis_tum_allcl_df) <- gtex_tissues_fin
corbytis_tum_allcl_df$`salivary gland` <- NULL # no salivary gland tumor samples
corbytis_tum_allcl_df$`small intestine` <- NULL # same reason
corbytis_tum_allcl_df$`nerve` <- NULL # same reason

corbytis_con_allcl_df <- tisgenesetcor(cl_tpm, con_tissue_tpm, tissue = gtex_tissues_fin[1], onlymatch = FALSE)
for(i in gtex_tissues_fin[-1]){
  corbytis_con_allcl_df <- cbind(corbytis_con_allcl_df, tisgenesetcor(cl_tpm, con_tissue_tpm, tissue = i, onlymatch = FALSE))
}
colnames(corbytis_con_allcl_df) <- gtex_tissues_fin
```

## make lil reference table for which cell lines came from what tissue/what their common name is
```{r}
master_sub <- sample_master_meta[sample_master_meta$rse_sampname %in% rownames(corbytis_tum_allcl_df),]
master_sub <- data.frame(SRA = master_sub$rse_sampname, tissue = master_sub$tissue_coerced, common_name = substr(master_sub$sample_name, 6, nchar(master_sub$sample_name)), lowercase_name = master_sub$cell_line)
# probably won't show their names necessarily but to reference
# did double check that cell lines are in same order as cor dfs
```

## generate heatmap
```{r}
library(pheatmap)

anno_col <- data.frame(master_sub$tissue)
rownames(anno_col) <- master_sub$SRA
colnames(anno_col) <- "Model Origin Tissue"

pheatmap(t(corbytis_tum_allcl_df), annotation_col = anno_col, show_colnames = FALSE, main = "Cell Line Median Correlation to Tumor Tissue by TAGs")

pheatmap(t(corbytis_con_allcl_df), annotation_col = anno_col, show_colnames = FALSE, main = "Cell Line Median Correlation to Control Tissue by TAGs")
```

## heatmap with top 100 cell lines
```{r}
# which cell lines have the highest median?
tum_top100cl <- data.frame(medcor = MatrixGenerics::rowMedians(as.matrix(corbytis_tum_allcl_df)), sra = rownames(corbytis_tum_allcl_df))
tum_top100cl <- tum_top100cl[order(tum_top100cl$medcor, decreasing = TRUE),][1:100,]$sra

con_top100cl <- data.frame(medcor = MatrixGenerics::rowMedians(as.matrix(corbytis_con_allcl_df)), sra = rownames(corbytis_con_allcl_df))
con_top100cl <- con_top100cl[order(con_top100cl$medcor, decreasing = TRUE),][1:100,]$sra

# subset cor dfs to top 100
corbytis_tum_top100cl_df <- corbytis_tum_allcl_df[match(tum_top100cl, rownames(corbytis_tum_allcl_df)),]
corbytis_con_top100cl_df <- corbytis_con_allcl_df[match(con_top100cl, rownames(corbytis_con_allcl_df)),]

# make anno cols, will be different now for tum/con because they have different cl's
tum_top100cl_annorow <- data.frame(Tissue = anno_col[match(tum_top100cl, rownames(anno_col)),]) # using match() to make sure it's in the order of tum_top100cl
rownames(tum_top100cl_annorow) <- tum_top100cl

con_top100cl_annorow <- data.frame("Origin Tissue" = anno_col[match(con_top100cl, rownames(anno_col)),]) # using match() to make sure it's in the order of con_top100cl
rownames(con_top100cl_annorow) <- con_top100cl

# I want to display the cell line common name
tum_top100_cl_names <- master_sub[match(tum_top100cl, master_sub$SRA),]$common_name
con_top100_cl_names <- master_sub[match(con_top100cl, master_sub$SRA),]$common_name

pheatmap(corbytis_tum_top100cl_df, annotation_row = tum_top100cl_annorow, labels_row = tum_top100_cl_names, fontsize_row = 6, main = "Top 100 Cell Line Median Correlation to Tumor Tissue by TAGs")

pheatmap(corbytis_con_top100cl_df, annotation_row = con_top100cl_annorow, labels_row = con_top100_cl_names, fontsize_row = 6, main = "Top 100 Cell Line Median Correlation to Control Tissue by TAGs")
```

### try heatmap with both tum and con tissue
```{r}
# top 100 cl's with highest median in BOTH
tum_topcl <- data.frame(medcor = MatrixGenerics::rowMedians(as.matrix(corbytis_tum_allcl_df)), sra = rownames(corbytis_tum_allcl_df))
con_topcl <- data.frame(medcor = MatrixGenerics::rowMedians(as.matrix(corbytis_con_allcl_df)), sra = rownames(corbytis_con_allcl_df))

mix_topcl <- cbind(tum_topcl, con_topcl[match(tum_topcl$sra, con_topcl$sra),])
mix_topcl <- data.frame(tumcor = mix_topcl[,1], concor = mix_topcl[,3], sra = mix_topcl[,2])
mix_topcl$meancor <- rowMeans(cbind(mix_topcl$tumcor, mix_topcl$concor))
mix_topcl <- mix_topcl[order(mix_topcl$meancor, decreasing = TRUE),]
mix_top100cl <- mix_topcl$sra[1:100]

# make new correlation matrix with both tissue types
corbytis_alltis_allcl_df <- cbind(corbytis_tum_allcl_df, corbytis_con_allcl_df)
colnames(corbytis_alltis_allcl_df) <- c(paste(stringr::str_to_title(colnames(corbytis_tum_allcl_df)), "Tumor"), paste("Non-Disease", stringr::str_to_title(colnames(corbytis_con_allcl_df))))

corbytis_alltis_100cl_df <- corbytis_alltis_allcl_df[rownames(corbytis_alltis_allcl_df) %in% mix_top100cl,]
corbytis_alltis_100cl_df <- corbytis_alltis_100cl_df[match(mix_top100cl, rownames(corbytis_alltis_100cl_df)),]

# make annotation row and column
## anno row = specifies origin tissue for each cell line
alltis_annorow <- cbind(anno_col, rownames(anno_col)) # to retain sra IDs for lining up
alltis_annorow <- alltis_annorow[rownames(alltis_annorow) %in% mix_top100cl,]
alltis_annorow <- alltis_annorow[match(mix_top100cl, alltis_annorow$`rownames(anno_col)`),]
alltis_annorow$`rownames(anno_col)` <- NULL
colnames(alltis_annorow) <- "Cell Line Tumor Origin "

## anno col = specifies whether tissue is control or tum
alltis_annocol <- data.frame("Tissue Type" = c(rep("Tumor", length(colnames(corbytis_tum_allcl_df))), rep("Non-Disease", length(colnames(corbytis_con_allcl_df)))))
rownames(alltis_annocol) <- colnames(corbytis_alltis_100cl_df)
colnames(alltis_annocol) <- "Tissue Type"

# cell line names
alltis_top100cl_names <- master_sub[match(mix_top100cl, master_sub$SRA),]$common_name

# heatmap!
ann_colors <- list("Tissue Type" = c('Non-Disease' = "darkblue", Tumor = "orange")) # make tissue type anno color consistent
pheatmap(t(corbytis_alltis_100cl_df), annotation_col = alltis_annorow, labels_col = alltis_top100cl_names, fontsize_col = 6, annotation_row = alltis_annocol, annotation_colors = ann_colors, fontface = "bold")

```