---
title: "TumPurity"
output: html_document
date: '2022-09-06'
---

> generating tumor purity info based on TCGA to determine if tumor purity highly influences model performance

# required packages
```{r}
library(readr)
library(foreach)
library(doParallel)
library(tibble)
```

# purity information from: https://gdc.cancer.gov/about-data/publications/pancanatlas ABSOLUTE purity/ploidy file (220906)
# code for computation sourced then adjusted from: https://github.com/katharineyu/TCGA_CCLE_paper/blob/master/R_scripts/tumor_cell_line_cor.R

# load in data from pancanatlas
```{r load}
purity_info_pca <- read.csv("../data/TCGA_mastercalls.abs_tables_JSedit.fixed.csv")
```

# load in TCGA RSE's with parallelized for loop
```{r}
rsefiles <- Sys.glob("../data/recount3/rse/tcga_tissue/*_tcga_rse.rds")

#setup for running in parallel
cl <- makeCluster(4)
registerDoParallel(cl, cores = 4)
  
tcga_barcode <- foreach(i = 1:length(rsefiles), .combine= c) %dopar% { #run loops in parallel, outputs a combined list
    rse <- readRDS(rsefiles[i])
    #raw <-  rse@assays@data@listData[["raw_counts"]]
    barcode <- strtrim(rse@colData@listData[["tcga.tcga_barcode"]],15)
}
stopCluster(cl) 
  
length(tcga_barcode[tcga_barcode %in% purity_info_pca$array]) # 9759 represented, 336 not
length(setdiff(tcga_barcode, purity_info_pca$array)) # 9759 represented, 336 not
```

# find highly tumor purity-associated genes
```{r computation}
#read in TCGA TPM
TCGA <- readRDS("../data/recount3/tpm/tcga_tpm.rds") 
#TCGA <- tibble::column_to_rownames(TCGA, var = "...1")
colnames(TCGA) <- tcga_barcode

#find TCGA samples with purity information
TCGA <- TCGA[, intersect(purity_info_pca$array, colnames(TCGA))]
tumors_purity <-  purity_info_pca[purity_info_pca$array %in% colnames(TCGA), ]
gc()

tmp <- lapply(rownames(TCGA), function(x)cor.test(as.numeric(TCGA[x,]),tumors_purity$purity, method = "s", exact = FALSE)) # returned lots of warnings 'Warning in cor.test.default(as.numeric(TCGA[x, ]), tumors_purity$purity,  :
  # Cannot compute exact p-value with ties'
# to overcome warnings, use exact = FALSE:  https://www.tutorialspoint.com/how-to-avoid-the-warning-cannot-compute-exact-p-value-with-ties-while-perform-correlation-test-for-spearman-s-correlation-in-r#:~:text=Since%20the%20spearman%20correlation%20coefficient,test%20function.
gc()

tmp.df <- data.frame(sapply(tmp,function(x)x$p.value), sapply(tmp,function(x)x$estimate))
rownames(tmp.df) <- rownames(TCGA)
colnames(tmp.df) <- c("pvalue", "rho")
tmp.df$padj <- p.adjust(tmp.df$pvalue)
tmp.df.sig <- tmp.df[tmp.df$padj < 0.01 & tmp.df$rho < -0.4 & complete.cases(tmp.df),]
purity_genes <- rownames(tmp.df.sig)
TCGA <- TCGA[!(rownames(TCGA) %in% purity_genes),]
```

# save tumor purity-associated genes
```{r save}
saveRDS(purity_genes, "../data/purity_genes.rds")
```


# clean env
```{r clean env}
rm(purity_info_pca, tcga_barcode, TCGA, tumors_purity, tmp, tmp.df, tmp.df.sig, purity_genes, tcga_rse)
gc()
```